RCloud: Integrating Exploratory Visualization, Analysis and Deployment

### 2014-03-26 discussion.tex
yes

Experience with deploying RCloud in a community of data analysts
has provided some insight into how its capabilities were received
that may be informative to researchers and designers of future systems
supporting visual analytics.
Returning to the design considerations outlined by Heer and Agrawala,
our experience underscores the value of several of their proposals,
and suggests areas for further exploration.

{\bf Shared artifacts and artifact histories} are a central feature.
The prototype aims at making it convenient to collaborate on code
and experiments with data. It was a success to observe that analysts
readily shared ideas through RCloud. We observed them creating experiments
in workbooks and sharing URLs to discuss data analysis and programming
problems, instead of sending screenshots in Powerpoint.

{\bf View sharing, bookmarking.} Likewise, everything is named
and accessed as a URL.

{\bf Discussion.} Annotation is supported in our human interface,
and was also a central goal for our prototype.
An interesting question is to decide to what extent the users of
production notebooks also use tools like annotation etc. without coding
or without being exposed to the analyst's view.\todo{Is this a gap in
our design?}

{\bf Content export} is not a capability we worked to support,
but the R environment already has many packages for it.
Our goal is to devise ways of working where explicit export is
less necessary. However it is desirable to integrate results
through web protocols such as JSON. We did assume that RCloud
notebooks need to play well in the ecosystem of other tools
and this worked out as planned, though it also leads to a temptation
to absorb features almost without limit. (Should RCloud support
templates for personal web pages? Should it incorporate emacs
on the web, and an email client?)

{\bf Social-psychological incentives} and {\bf voting and ranking}
are supported through starring. This mechanism was employed often
by our test audience. An obvious next step would be to enhance
recommendations using relationships discovered by static and
dynamic code analysis. This should be done both within and across
collections of scripts (the latter being similar to work in Vistrails
that enhanced recommendations by clustering multiple workflows).
It may be valuable to data analysts to know which packages are often
used together, or with a given type of data. It is probably helpful
to know which data sources and packages often go together.

{\bf Group management, size and diversity.} This area lacks sufficient
support in RCloud, but is clearly desirable. We rely on conventional
administrative and social processes, outside our platform. Automation
of this capability in a large organization could rely on social media
services

@@

>>>>

Experience with deploying RCloud in a community of data analysts
has provided some insight into how well the proposed requirements
satisfied their needs. This may be informative to researchers and
designers of future systems supporting visual analytics.

Returning to the design considerations outlined by Heer and Agrawala,
our experience underscores the value of several of their proposals,
and suggests areas for further exploration.

{\bf Shared artifacts and artifact histories} are a central feature.
The prototype aimed at making it convenient to collaborate on code
and experiments with data. It was a success to observe that analysts
readily shared ideas through RCloud. We observed them creating experiments
in workbooks and sharing URLs to discuss analysis and programming
problems, instead of sending screenshots in Powerpoint.

{\bf View sharing, bookmarking.} Likewise, everything is named
and accessed as a URL. This turned out to be an effective mechanism
to support sharing and integration of analyses in notebooks
with external processes and systems.

{\bf Discussion.} Annotation is supported in our human interface,
and was also a central goal for our prototype. \todo{Elaborate.}
An interesting question is to what extent users of production notebooks
should be able to use tools like annotation without coding and
without being exposed to the analyst's view.\todo{Is this a gap in
our design?}

{\bf Content export} is not a capability we worked to support,
but the R environment already has many packages for this.
Our goal was to devise ways of working where explicit export is
less necessary. However it is desirable to integrate results
through web protocols such as JSON. We did assume that RCloud
notebooks need to play well in the ecosystem of other tools,
and this worked out as intended, though we learned it also creates
a temptation to absorb features almost without limit. (Should
RCloud support templates for personal web pages? Should it
incorporate emacs on the web, and an email client?)

{\bf Social-psychological incentives} and {\bf voting and ranking}
are supported through starring. This mechanism was employed often
by our test audience.  An obvious next step would be to enhance
recommendations using relationships discovered by static and
dynamic code analysis. This should be done both within and across
collections of scripts (the latter being similar to work in Vistrails
that enhanced recommendations by clustering multiple workflows).
It may be valuable to data analysts to know which packages are often
used together, or with a given type of data. It is probably helpful
to know which data sources and packages often go together. In general,
though, trivial or even passive mechanisms to collect data for
recommendations are essential.

{\bf Group management, size and diversity.} This area lacks sufficient
support in RCloud, but is clearly desirable. We rely on conventional
external administrative and social processes, outside our platform,
to manage accounts and groups.  Automation of this capability in a
large organization could rely on social media services

@@

Our environment assumes a high level of trust within the user community.
Access control for information security is implemented by relying on the
host operating system and web server. We admit this may require further
